### 2013-03-30 exam6/2.tex
outline the problems

JSON is a lightweight data-interchange format.  It is widely used on
the web and plays exactly the role of S-Expressions in universe
programs, but for JavaScript programs.  It's used by folks like Google
and there's tons of data on the web that's available in the JSON
format.  So as to not be left out of the fun, your co-op employer has
asked you to design an object representation of JSON to incorporate
into their 2 million lines of {\tt class/2} code.  To start things off
simply, here are some examples of what we'll call ``JSON, Jr.'':

\begin{verbatim}
"this is JSON"
[ "this is JSON", "Jr." ]
[ [], [ "So", "is" ], "this" ]
\end{verbatim}
The first example is just a string.  The second is an ``array'' of two
JSON elements, which are both strings.  The third is another array,
but of three elements: the first is an array of zero elements, the
second is an array of two strings, and the third is a string.

\begin{enumerate}
\item Design an object-based data representation for JSON, Jr.

\ifrubric
\begin{alltt}
;; A JSON is one of:  [5pt]
;; - (new s% String)
;; - Array
;;
;; An Array is one of
;; - (new a0%)
;; - (new a+% JSON Array)

(define-class s%
  (fields val))

(define-class a0%)

(define-class a+%
  (fields hd tl))
\end{alltt}
\else
\newpage
[Here is some more space for the previous problem.]
\fi
\newpage

\item Design a method for counting the number of strings in a JSON,
  Jr.~object.
\ifrubric
\begin{alltt}
;; JSON implements [1pts]
;; - Count the number of strings in this object
;;   count-strings : -> Natural

;; in s% [1pt]
(define (count-strings) 1)

;; in a0% [1pt]
(define (count-strings) 0)

;; in a+% [1pt]
(define (count-strings)
  (+ (this . hd . count-strings)
     (this . tl . count-strings)))

;; [1pts]
(define eg1 (new s% "This is JSON"))
(define eg2 (new a+% eg1 (new a+% (new s% "Jr.") (new a0%))))
(define eg3 (new a+% (new a0%)
                 (new a+% (new a+% (new s% "So")
                               (new a+% (new s% "is") (new a0%)))
                      (new a+% (new s% "this")
                           (new a0%)))))

;; [1pts]
(check-expect (eg1 . count-strings) 1)
(check-expect (eg2 . count-strings) 2)
(check-expect (eg3 . count-strings) 3)
\end{alltt}
\else
\fi
\newpage
\item Among the 2M lines of code your employer maintains are 30,000
  lines of code devoted to implementations of the {\tt
    StringPred} interface, which is:
\begin{verbatim}
;; A StringPred implements
;; - apply : String -> Boolean
\end{verbatim}
These objects represent predicates on strings.  Now that your JSON,
Jr. library is in place, they've decided to put their predicates to
use in searching a large corpus of JSON data.  To accomodate this,
design the following method for JSON, Jr.~objects:
\begin{verbatim}
;; Find the first string in this JSON Jr object satisfying
;; given predicate, or #f if there's no such string.
;; find : StringPred -> String or #f
\end{verbatim}
\ifrubric
\begin{alltt}
;; in s%  [2pts]
(define (find p)
  (if (p . apply (this . val))
      (this . val)
      #f))

;; in a0% [1pts]
(define (find p) false)

;; in a+% [3pts]
(define (find p)
  (local [(define r (this . hd . find p))]
    (if (false? r)
        (this . tl . find p)
        r)))

;; [2pts]
;; A (new eq% String) implements StringPred
(define-class eq%
  (fields x)
  (define (apply y)
    (string=? (this . x) y)))

;; [1pts]
(check-expect (eg1 . find (new eq% "This is JSON")) "This is JSON")
(check-expect (eg1 . find (new eq% "this")) false)
(check-expect (eg2 . find (new eq% "This is JSON")) "This is JSON")
(check-expect (eg2 . find (new eq% "this")) false)
(check-expect (eg3 . find (new eq% "This is JSON")) false)
(check-expect (eg3 . find (new eq% "this")) "this")
\end{alltt}
\else
\fi
\end{enumerate}
\end{problem}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{problem} \pts{19}

% Set/Property problem

In order to manage the distribution of ID cards, the University needs
a data structure for managing sets of numbers.  Van Horn started on
the implementation, resulting in what you see below.

\begin{verbatim}
;; A Set is one of:
;; - (new empty%)
;; - (new singelton% Number)
;; - (new union% Set Set)
;; Interpretation: a set that contains all the *unique* numbers in the data
;; and implements
;; - sum : -> Number
;;   Produce the sum of all the unique elements of this set
;; - sum/acc : Set -> Number
;;   Produce the sum of all the unique elements of this set that are
;;   *not* in the given set
;; - contains : Number -> Boolean
;;   Does this set contain the given number?
;; - union : Set -> Set
;;   Combine this set with the given set

(define-class empty%
  (define (sum) 0)
  (define (sum/acc s) 0))

(define-class singleton%
  (fields n)
  (define (sum) (this . n))
  (define (sum/acc s) (cond [(s . contains (this . n)) 0]
                            [else (this . n)])))

(define-class union%
  (fields l r)
  (define (sum) (this . sum/acc (new empty%)))
  (define (sum/acc s) (+ (this . l . sum/acc s)
                         (this . r . sum/acc (s . union (this . l))))))
\end{verbatim}

\newpage

Unfortunately, as you can see, Van Horn fell asleep before finishing,
and the deadline is today.  Help him out by implementing the
\verb|union| and \verb|contains| methods for all \verb|Set|s.

\ifrubric
\begin{alltt}
\rubricpts{1 pt -- union method}
\rubricpts{3 pts -- contains method}
;; in empty%
(define (union s) s)
(define (contains n) #f)
;; also ok:
(define (union s) (new union% this s))
;; in singleton%
(define (union s) (new union% this s))
(define (contains n) (= n (this . n)))
;; in union%
(define (union s) (new union% this s))
(define (contains n) (or (this . l . contains n)
                         (this . r . contains n)))
(define e (new empty%))
\rubricpts{2 pts -- tests}
(check-expect (e . contains 5) #f)
(check-expect ((new singleton% 5) . contains 5) #t)
(check-expect ((new singleton% 6) . contains 5) #f)
(check-expect ((new singleton% 6) . union  e . contains 5) #f)
(check-expect (e . union (new singleton% 5) . contains 5) #t)
(check-expect ((new singleton% 5) . union (new singleton% 5) . contains 5) #t)
\end{alltt}
\else
\newpage
[Here is some more space for the previous problem.]
\fi
\newpage

\verb|Set|s should satisfy the following property, called the
``union-contains'' property, for any number $n$ and
any two sets $s_1$ and $s_2$:

\begin{quote}
Given two sets $s_1$ and $s_2$, if $n$ is contained in the union of $s_1$ and $s_2$, then $n$ is either
contained in $s_1$ or it is contained in $s_2$.
\end{quote}

\begin{enumerate}
\item Codify the ``union-contains'' property as a predicate.  Write a
  test case that should pass (if your implementations work right) using the
  predicate you have defined.
\end{enumerate}

\ifrubric
\begin{alltt}
\rubricpts{1 pt -- contract & purpose}
;; union-contains : Set Set Number -> Boolean
\rubricpts{3 pts -- code}
(define (union-contains s1 s2 n)
  (check-expect (s1 . union s2 . contains n)
                (or (s1 . contains n) (s2 . contains n))))
\rubricpts{1 pt -- successful test}
(union-contains e (new singleton% 17) 17)
\end{alltt}
\else
\newpage
[Here is some more space for the previous problem.]
\fi

\newpage
\noindent
The University doesn't really trust Van Horn, and they want to make
sure that \verb|Set|s satisfy another property, called the
``union-sum'' property:

\begin{quote}
Given two sets $s_1$ and $s_2$, computing the sum of $s_1$ and the sum
of $s_2$ and then adding the two results is equal to the sum of the
union of  $s_1$ and $s_2$.
\end{quote}

\noindent
The University is considering adding this to the test suite as well.

\begin{enumerate}
\setcounter{enumi}{1}

\item Codify this property as a predicate, and write a test case that
  should pass using the predicate.

\item Is the ``union-sum'' property true for all sets?  If so,
  give an explanation.  If not, write a counter-example (i.e., a
  failing test case) using your predicate.

\end{enumerate}

\begin{solution}
\begin{alltt}
\rubricpts{1 pt -- contract & purpose}
;; union-sum : Set Set -> Boolean
\rubricpts{3 pts -- code}
(define (union-sum s1 s2 n)
  (check-expect (s1 . union s2 . sum)
                (+ (s1 . sum) (s2 . sum))))

\rubricpts{1 pt -- successful test}
;; successful test
(union-sum (new singleton% 17) e)
\rubricpts{3 pts -- failing test}
;; It's false, counterexample:
(union-sum (new singleton% 17) (new singleton% 17))
\end{alltt}
\end{solution}

\newpage
\ifrubric\else
[Here is some more space for the previous problem.]
\newpage
\fi
\end{problem}

@@

\newpage

@@

\begin{problem}\pts{5}

% Visitor on Binary Trees

\textbf{Extra Credit}
\\
\noindent
Labich was unhappy to discover that the class languages don't have his
favorite programming language feature---pattern matching.  After
thinking hard about this problem, he came up with a solution, which he
calls \emph{pattern objects}.

Consider the following data defintion:
\begin{verbatim}
;; A BTree is one of:
;; - (new leaf% Number)
;; - (new node% BTree Number BTree)
;; and implements
;; - match : [Pattern X] -> X
;;   applies the given pattern object to this tree

;; A [Pattern X] implements:
;; - leaf : Number -> X
;; - node : BTree Number BTree -> X

(define-class leaf%
  (fields n)
  ;; apply the appropriate method from the given pattern object
  (define (match p) (p . leaf (this . n))))

(define-class node%
  (fields left n right)
  ;; apply the appropriate method from the given pattern object
  (define (match p) 
    (p . node (this . left) (this . n) (this . right))))
\end{verbatim}

A pattern object is like the implementation of two different functions
at once: one that implements a function for leaves from a binary tree,
and one function that works on nodes.

Write a pattern object that computes the product of all of the numbers
in a binary tree.

\newpage
\ifrubric\else
[Here is some more space for the previous problem.]
\fi

\begin{solution}
\begin{alltt}
\rubricpts{5 pts -- get it right}
;; product% implements [Pattern Number]
(define-class product%
  ;; the product of a leaf
  (define (leaf n) n)
  ;; the product of a node
  (define (node l n r) (* n (l . match this) (r . match this))))

(define p (new product%))
(check-expect ((new leaf% 5) . match p) 5)
(check-expect ((new node%  (new leaf% 2) 5 (new leaf% 7)) . match p) 70)
\end{alltt}
\end{solution}
\end{problem}
\end{document}

>>>>

 Binary tree problem

\ifrubric
\else
\newpage
[Here is some more space for the previous problem.]
\fi
\newpage
\newpage

\textbf{Extra Credit}

\end{problem}

@@

@@

Outline problem

\ifrubric
\else
\newpage
[Here is some more space for the previous problem.]
\fi
\newpage
\newpage

\end{problem}

@@

Computation/thunk problem

\ifrubric
\else
\newpage
[Here is some more space for the previous problem.]
\fi
\newpage
\newpage

\end{problem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
